Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-31 article/02_Materials_and_Methods.Rmd
reorganized general file structure

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
\caption{\label{map_graves}All graves dating between 2200 and 800 calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}). The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape. The rectangular frame indicates the research area. The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from Natural Earth project (\url{www.naturalearthdata.com}).}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../analysis/figures/map_regions.jpeg}
\caption{\label{map_regions}Circular, artificial regions (\textbf{A}) and spatial distance network with classified distance values (\textbf{B}). The map layout is similar to Figure \ref{map_graves} but instead of rivers modern country borders are drawn.}
\end{figure}

@@

```{r gprcrosstab, echo=FALSE, tidy=FALSE}
load("../analysis/data/output_data/gprcrosstab.RData")
colnames(gprcrosstab)[1:2] <- c("Region", "Burial type")
gprcrosstab %>% 
  knitr::kable(
    caption = "Burial rite distribution in the active grave dataset. The order of the regions follows a southeast-northwest direction.",
    format = "latex",
    booktabs = T,
    escape = F,
    align = "l"
  ) %>%
  kableExtra::kable_styling(
    full_width = T,
    font_size = 9,
    latex_options = "hold_position"
  ) %>%
  kableExtra::add_header_above(c(" " = 1, " " = 1, "Burial construction" = 3)) %>%
  kableExtra::column_spec(1, width = "2 cm") %>%
  kableExtra::column_spec(2, width = "1.5 cm") %>%
  kableExtra::column_spec(3:5, width = "0.8cm") %>%
  kableExtra::collapse_rows(1, row_group_label_position = "identity", valign = "top") %>%   kableExtra::row_spec(row = 25:27, align = "center", bold = T)

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../analysis/figures/development_burial_type.jpeg}
\caption{\label{development_burial_type}Regional distribution time series of the variable \textit{burial type} with the absolute amount (\textbf{A}) and proportion (\textbf{B}) of grave types. The x-axis shows the time, the y-axis the amount or the percentage shares of graves. The individual plots are separated by region, and each region is featured by a small map according to Figure \ref{map_regions} A. $n$ in the facet labels equals the total number of available dates with information for the variable (cf. Table \ref{tab:gprcrosstab}).}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../analysis/figures/development_burial_construction.jpeg}
\caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.\newline\newline\newline\newline\newline}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{../analysis/figures/sed_region_matrix.jpeg}
\caption{\label{sed_region_matrix}Plot matrix of the region-region squared Euclidean distance time series. In the top right half with solid lines for the variable \textit{burial type} and in the bottom left with dashed lines for the variable \textit{burial construction}. The x-axis displays time, the y-axis the squared Euclidean distance between two relevant regions. The faint line shows the actual data, the stronger line an approximation (LOESS with $span = 0.3$).}
\end{figure*}

@@

```{r dacta, echo=FALSE, tidy=FALSE}
load("../analysis/data/output_data/distance_and_correlation_table_A.RData")
distance_and_correlation_table_A %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Mean|Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black"))) 
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1)) 
  ) %>%
knitr::kable(
  "latex", 
  caption = "Mean distance, correlation and mantel test results for cultural and spatial distance. Listed are the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B). $p$-values below $0.1$ are printed bold.", 
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "burial type & spatial distance" = 3, "burial construction & spatial distance" = 3)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "A" = 3, "B" = 3),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

```{r dactb, echo=FALSE, tidy=FALSE}
load("../analysis/data/output_data/distance_and_correlation_table_B.RData")
distance_and_correlation_table_B %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black")))
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1))
  ) %>%
knitr::kable(
  "latex",
  caption = "Like Table \\ref{tab:dacta}. \\textbf{C} shows the Mantel test results for the two variables if compared directly, \\textbf{D} the results of the Partial Mantel test considering spatial distance as a second independent variable.",
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "variable correlation" = 2, "variable correlation with spatial correction" = 2)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "C" = 2, "D" = 2),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../analysis/figures/correlation_time_series.jpeg}
\caption{\label{simulation}Development of correlation between cultural and spatial distance. The results for the burial customs data (Table \ref{tab:dacta} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{map_graves.jpeg}
\caption{\label{map_graves}All graves dating between 2200 and 800 calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}). The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape. The rectangular frame indicates the research area. The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from Natural Earth project (\url{www.naturalearthdata.com}).}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{map_regions.jpeg}
\caption{\label{map_regions}Circular, artificial regions (\textbf{A}) and spatial distance network with classified distance values (\textbf{B}). The map layout is similar to Figure \ref{map_graves} but instead of rivers modern country borders are drawn.}
\end{figure}

@@

```{r gprcrosstab, echo=FALSE, tidy=FALSE}
load("gprcrosstab.RData")
colnames(gprcrosstab)[1:2] <- c("Region", "Burial type")
gprcrosstab %>% 
  knitr::kable(
    caption = "Burial rite distribution in the active grave dataset. The order of the regions follows a southeast-northwest direction.",
    format = "latex",
    booktabs = T,
    escape = F,
    align = "l"
  ) %>%
  kableExtra::kable_styling(
    full_width = T,
    font_size = 9,
    latex_options = "hold_position"
  ) %>%
  kableExtra::add_header_above(c(" " = 1, " " = 1, "Burial construction" = 3)) %>%
  kableExtra::column_spec(1, width = "2 cm") %>%
  kableExtra::column_spec(2, width = "1.5 cm") %>%
  kableExtra::column_spec(3:5, width = "0.8cm") %>%
  kableExtra::collapse_rows(1, row_group_label_position = "identity", valign = "top") %>%   kableExtra::row_spec(row = 25:27, align = "center", bold = T)

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{development_burial_type.jpeg}
\caption{\label{development_burial_type}Regional distribution time series of the variable \textit{burial type} with the absolute amount (\textbf{A}) and proportion (\textbf{B}) of grave types. The x-axis shows the time, the y-axis the amount or the percentage shares of graves. The individual plots are separated by region, and each region is featured by a small map according to Figure \ref{map_regions} A. $n$ in the facet labels equals the total number of available dates with information for the variable (cf. Table \ref{tab:gprcrosstab}).}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{development_burial_construction.jpeg}
\caption{\label{development_burial_construction}Like Figure \ref{development_burial_type}, but for the variable \textit{burial construction}.\newline\newline\newline\newline\newline}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{sed_region_matrix.jpeg}
\caption{\label{sed_region_matrix}Plot matrix of the region-region squared Euclidean distance time series. In the top right half with solid lines for the variable \textit{burial type} and in the bottom left with dashed lines for the variable \textit{burial construction}. The x-axis displays time, the y-axis the squared Euclidean distance between two relevant regions. The faint line shows the actual data, the stronger line an approximation (LOESS with $span = 0.3$).}
\end{figure*}

@@

```{r dacta, echo=FALSE, tidy=FALSE}
load("distance_and_correlation_table_A.RData")
distance_and_correlation_table_A %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Mean|Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black"))) 
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1)) 
  ) %>%
knitr::kable(
  "latex", 
  caption = "Mean distance, correlation and mantel test results for cultural and spatial distance. Listed are the overall mean squared Euclidian distances for each 200 years time slot of all region-region relations and the result statistics of the Mantel test for the variables \\textit{burial type (bt)} (\\textbf{A}) and \\textit{burial construction (bc)} (\\textbf{B}) with the spatial (sp) distance matrix (Figure \\ref{map_regions} B). $p$-values below $0.1$ are printed bold.", 
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "burial type & spatial distance" = 3, "burial construction & spatial distance" = 3)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "A" = 3, "B" = 3),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

```{r dactb, echo=FALSE, tidy=FALSE}
load("distance_and_correlation_table_B.RData")
distance_and_correlation_table_B %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("Correlation")),
    list(~kableExtra::cell_spec(., "latex", color = ifelse(. < 0, "red", "black")))
  ) %>%
  dplyr::mutate_at(
    dplyr::vars(dplyr::matches("p-Value")),
    list(~kableExtra::cell_spec(., "latex", bold = . < 0.1))
  ) %>%
knitr::kable(
  "latex",
  caption = "Like Table \\ref{tab:dacta}. \\textbf{C} shows the Mantel test results for the two variables if compared directly, \\textbf{D} the results of the Partial Mantel test considering spatial distance as a second independent variable.",
  booktabs = T,
  linesep = "\\addlinespace \\hline \\addlinespace",
  escape = F
) %>%
  kableExtra::add_header_above(
    c(" ", "variable correlation" = 2, "variable correlation with spatial correction" = 2)
  ) %>%
  kableExtra::add_header_above(
    c(" ", "C" = 2, "D" = 2),
    bold = T, line = F
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9,
    full_width = T
  ) %>%
  kableExtra::column_spec(c(1), width = "4cm") %>%
  gsub("table", "table*", .)
```

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
\caption{\label{simulation}Development of correlation between cultural and spatial distance. The results for the burial customs data (Table \ref{tab:dacta} AB) are displayed with a solid (\textit{burial type}) and dashed (\textit{burial construction}) line. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed.}
\end{figure}

### 2019-03-31 article/index.Rmd
reorganized general file structure

---
title: Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data
runninghead: Schmid.
author:
- name: Clemens Schmid
  num: 1
address:
- num: 1
  org: Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, Mainz, Germany
corrauth: Clemens Schmid, Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, D - 55116 Mainz.
email: clemens@nevrome.de
abstract: European Bronze Age archaeology traditionally focusses on two major dimensions to categorise burials -- although there is an immense variability of attendant phenomena within this spectrum -- flat graves versus burial mounds and cremation versus inhumation. These traits are an indispensable ingredient for common archaeological narratives of sociocultural interaction and cultural evolution. This paper presents a quantitative reconstruction of the general trends in the distribution of Bronze Age burial traditions based on bulk radiocarbon data and employs the resulting time series for the estimation of macro-regional cultural distance. Despite the relatively small amount of input data the trend reconstruction fits to established archaeological observations for prehistoric Europe. The comparison of cultural and spatial distance leads to the remarkable result of no significant permanent correlation, which indicates that the spread of the relevant aspects of burial traditions can not be explained with simple diffusion models. Instead a more complex process of cultural transmission has to be considered.
keywords: Bronze Age, Burial Traditions, Cultural Transmission, Radiocarbon Dating 
classoption:
  - Afour
  - times
bibliography: ../references.bib
bibliographystyle: apalike
output:
  rticles::sage_article:
    keep_tex: yes
    includes:
        in_header: preamble.tex
---

@@

# allow easy dynamic value integration
sf_prep <- "../analysis/data/output_data/sf_prep.txt"
sf_desc <- "../analysis/data/output_data/sf_desc.txt"
r <- txtstorage::read
a <- function(x) {r("regions graves amounts", sf_prep) %>% strsplit(", ") %>% unlist %>% magrittr::extract(grepl(x, .))}
```

>>>>

---
title: Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data
runninghead: Schmid.
author:
- name: Clemens Schmid
  num: 1
address:
- num: 1
  org: Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, Mainz, Germany
corrauth: Clemens Schmid, Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, D - 55116 Mainz.
email: clemens@nevrome.de
abstract: European Bronze Age archaeology traditionally focusses on two major dimensions to categorise burials -- although there is an immense variability of attendant phenomena within this spectrum -- flat graves versus burial mounds and cremation versus inhumation. These traits are an indispensable ingredient for common archaeological narratives of sociocultural interaction and cultural evolution. This paper presents a quantitative reconstruction of the general trends in the distribution of Bronze Age burial traditions based on bulk radiocarbon data and employs the resulting time series for the estimation of macro-regional cultural distance. Despite the relatively small amount of input data the trend reconstruction fits to established archaeological observations for prehistoric Europe. The comparison of cultural and spatial distance leads to the remarkable result of no significant permanent correlation, which indicates that the spread of the relevant aspects of burial traditions can not be explained with simple diffusion models. Instead a more complex process of cultural transmission has to be considered.
keywords: Bronze Age, Burial Traditions, Cultural Transmission, Radiocarbon Dating 
classoption:
  - Afour
  - times
bibliography: references.bib
bibliographystyle: apalike
output:
  rticles::sage_article:
    keep_tex: yes
    includes:
        in_header: preamble.tex
---

@@

# allow easy dynamic value integration
sf_prep <- "sf_prep.txt"
sf_desc <- "sf_desc.txt"
r <- txtstorage::read
a <- function(x) {r("regions graves amounts", sf_prep) %>% strsplit(", ") %>% unlist %>% magrittr::extract(grepl(x, .))}
```
